Enhancing Accelerometer-based Activity Recognition with Capacitive Proximity Sensing

### 2012-06-19 Paper/Paper.tex
updated two images in the paper
updated two images
changed tex

[MB; EB: Introduction to HedgeHog device]

@@

image : photo

\begin{figure}[htbp]
	\centering
		\includegraphics[width=1.00\textwidth]{Images/schematic.pdf}
	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower.}
	\label{fig:schematic}
\end{figure}

@@

The experiment setup aims to depict a typical scenario in daily life. Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping. A fine-grained monitoring of such activities may help elderly or people suffering from mental diseases to maintain a healthy day/night rhythm and take action if irregularities occur.

@@

\begin{figure}
	\centering
		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/drinking.pdf}
	\caption{An example of the ``drinking'' activity with raw 3D acceleration data (lower plot) and proximity sensor data (upper plot). The participant took five nips from the glass, while leaving the hands lying on the table in between. These motions can be easily detected in the acceleration as well as proximity data.}
	\label{fig:drinking}
\end{figure}

\begin{figure}
	\centering
		\includegraphics[trim=1.9cm 0.5cm 1.8cm 0.5cm,clip,width=\textwidth]{Images/making-bread.pdf}
	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast. Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.}
	\label{fig:making-bread}
\end{figure}

@@

In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with ten test persons. All test persons received a plot with the activities they were supposed to perform. After manual labeling, we used this test-set as ground truth and performed a cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set.

[EB MB: kurze Einführung in SVMs]

The classifier was trained with various features that were extracted from windows with the length of 5 seconds. In order to suppress noise contained in the capacitive proximity sensing data, we applied a Gaussian kernel with a size of [TODO] samples. The final feature set contained the arithmetic mean and standard variance of both accelerometer and proximity data. Moreover, we included the median, min and max features for both modalities.

@@

\newpage

>>>>

@@

\begin{figure}[htbp]
	\centering
		\includegraphics[width=1.00\textwidth]{Images/schematic.pdf}
	\caption{The sensing circuit is based on a timer with an operational amplifier that acts as a voltage follower. The wire that is labeled with "`antenna"' leads to the sensing electrode, whereas the wire labeled with "`guard"' leads to the shield electrode.}
	\label{fig:schematic}
\end{figure}

@@

The experiment setup aims to depict a typical scenario in daily life. Especially in the field of Ambient Assisted Living (AAL), it is desired to monitor activities like drinking, preparing lunch and sleeping. A fine-grained monitoring of such activities may help elderly or people suffering from mental diseases to maintain a healthy day/night rhythm and take action if irregularities occur. Figure blubb shows the modified HedgeHog activity logger that has been extended with a capacitive proximity sensor. The wristband has two electrodes, a sensing electrode underneath a slightly bigger shield electrode.

[Hier Figure Aufbau]

@@

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_7.pdf}
	\caption{An example of the ``drinking'' activity with raw 3D acceleration data (lower plot) and proximity sensor data (upper plot). The participant took five nips from the glass, while leaving the hands lying on the table in between. These motions can be easily detected in the acceleration as well as proximity data.}
	\label{fig:drinking}
\end{figure}

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{../Auswertung/images/eugen_5.pdf}
	\caption{An example of the ``preparing bread'' activity, where the participants had to put marmalade on a toast. Note, that the acceleration data does not seem to have any characteristic patterns, while the proximity sensor indicates a table, plate, or other objects in immediate distance.}
	\label{fig:making-bread}
\end{figure}

@@

In order to evaluate if capacitive proximity sensors in wrist-bands can enhance the performance of activity recognition, we have conducted an evaluation with seven test persons. All test persons received a plot with the activities they were supposed to perform. After manual labeling, we used this test-set as ground truth and performed a 4-fold cross-validation on an support-vector machine (SVM) classifier once with and once without including the data of the capacitive proximity sensor into the feature set. We used an SVM classifier because of its high relevance in activity recognition and its fast performance.

The classifier was trained with various features that were extracted from windows with the length of 1 second. It turned out that greater window sizes do not provide better results. In order to suppress noise contained in the capacitive proximity sensing data, we applied a moving average filter with a kernel size of 10. The final feature set contained the arithmetic mean and standard variance of both accelerometer and proximity data. These simple feature types represent standard features applied in activity recognition.

@@

% \newpage
